---
title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
keep_md: true
---

### Loading and preprocessing the data

The activity data is loaded into R using read.csv() command. 

```{r read_data, echo=TRUE}

activity_data <- read.csv("activity.csv")
```

### What is mean total number of steps taken per day?

The total number of steps taken per day is calculated using aggregate().

```{r, echo=TRUE}
total_steps_per_day <- aggregate(steps~date, data=activity_data, FUN=sum)
```
The histogram for total steps per day is plotted, and the mean and median for total steps per day are reported.

```{r histogram, echo=TRUE}
hist(total_steps_per_day$steps, 
     main="Histogram of total number of steps per day",
     ylim = c(0,30),
     xlab="Total number of steps per day")
```

```{r, mean ane median for total steps per day, echo=TRUE}
mean(total_steps_per_day$steps)
median(total_steps_per_day$steps)
```

### What is the average daily activity pattern?

The average steps per interval are calculated using aggregate(). These values are plotted against the time intervals. The interval with the largest number of steps is retrieved using which() function. 
```{r, echo=TRUE}

average_steps_per_interval <- aggregate(steps ~ interval, 
                                        data=activity_data, 
                                        FUN=mean, na.rm=T)
#Plot the average steps against intervals
plot(steps ~ interval, data=average_steps_per_interval, type="l")

#Which interval has the largest average steps per day?
average_steps_per_interval$interval[which.max(average_steps_per_interval$steps)]

```

### Imputing missing values

The total number of data with NA in the dataset is calculated.

```{r total number of NAs, echo=TRUE}
sum(apply(activity_data, 1, function(x) any(is.na(x))))
```

In the following code, the missing values are being replaced by the mean for that 5-minute interval.

```{r imputing the missing values, echo=TRUE}
old <- activity_data
for (i in 1:nrow(activity_data)) {
        
        if (is.na(activity_data[i,"steps"])) {
                
                activity_data[i, "steps"] <- average_steps_per_interval$steps[average_steps_per_interval$interval == activity_data[i, "interval"]]
                
        }
}
activity_data_new <- activity_data
activity_data_old <- old

total_steps_per_day_new <- aggregate(steps~date, data=activity_data_new, FUN=sum)
hist(total_steps_per_day_new$steps, 
     main="Histogram of total number of steps per day",
     ylim = c(0,40),
     xlab="Total number of steps per day")

# mean of the total number of steps in the new dataset
mean(total_steps_per_day_new$steps)

# median for the new dataset
median(total_steps_per_day_new$steps)    
```
Compared with the old dataset, the mean did not change since NAs were replaced with means for for the same time interval. There was a slight change in the median.

### Are there differences in activity patterns between weekdays and weekends?

A new factor variable with two levels is added to the new dataset. Then, the mean number of steps for a time interval is calculated separately for each level. 

```{r weekdays vs weekends, echo = TRUE}
dates <- weekdays(as.Date(activity_data_new$date))
weekdays <- ifelse(dates%in%c("Saturday","Sunday"), "Weekend","Weekdays")
activity_data_new$weekdays <- weekdays
activity_data_new$weekdays <- as.factor(activity_data_new$weekdays)

weekdays_data <- activity_data_new[activity_data_new$weekdays=="Weekdays",]
weekend_data <- activity_data_new[activity_data_new$weekdays=="Weekend",]

average_steps_per_interval_weekdays <- aggregate(steps ~ interval, 
                                        data=weekdays_data, 
                                        FUN=mean, na.rm=T)

plot(steps ~ interval, data=average_steps_per_interval_weekdays, 
     type="l", col="blue", main="Mean steps per interval")

average_steps_per_interval_weekend <- aggregate(steps ~ interval, 
                                        data=weekend_data, 
                                        FUN=mean, na.rm=T)

lines(steps ~ interval, data=average_steps_per_interval_weekend, 
      type="l", col="red")
legend("topright", c("Weekdays","Weekend"), col=c("blue","red"), pch=19)  

```    
    
There is a difference in activity patterns between weekdays and weekends. The number of mean steps was higher in the time interval 500-1000 during weekdays, whereas it was higher in the time interval 1000-1700 during weekends.